% Define block styls
\tikzstyle{line} = [draw, very thick, color=black!50, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=2em]
\tikzstyle{decision} = [diamond, draw, fill=blue!20,
    text badly centered, node distance=6em, inner sep=0pt]
\tikzstyle{block} = [rectangle, draw, %fill=blue!20,
    text centered, node distance=4em]
\tikzstyle{iogram} = [trapezium, draw, fill=black!10, %fill=pink!20,
    trapezium left angle=70, trapezium right angle=-70,
    text centered, align=center, node distance=5em]

\DeclareRobustCommand\suppOrApp{%
  \ifthenelse{\boolean{thesisStyle}}
    {Appendix}
    {Supplementary}
}


\ifthenelse{\boolean{thesisStyle}}
{ 
% \renewcommand{\caption}[1]{\singlespacing\hangcaption{#1}\normalspacing}
  \renewcommand{\topfraction}{0.85}
  \renewcommand{\textfraction}{0.1}
%
% This seems to cause a problem with floats and excess vertical space:
%
  \renewcommand{\floatpagefraction}{0.75}
}
{}


%\newcommand{\ANDw}{\textnormal{AND}}
%\newcommand{\ORw}{\textnormal{OR}}
\newcommand{\ANDw}{\land}
\newcommand{\ORw}{\lor}

\newcommand{\falconAbstractMotivation}{%
A major theme in constraint-based modeling is unifying 
experimental data, such as biochemical information about the reactions
that can occur in a system or the composition and localization of enzyme
complexes, with high-throughput data including expression data,
metabolomics, or DNA sequencing. The desired result is to increase
 predictive capability resulting in improved understanding of metabolism.
 The approach typically employed when only gene (or protein) intensities
are available is the creation of tissue-specific models, which reduces
the available reactions in an organism model, and does not provide an
objective function for the estimation of fluxes, which is an important
limitation in many modeling applications.
}

\newcommand{\falconAbstractResults}{%
We develop a method, flux assignment with LAD (least absolute
deviation) convex objectives and normalization (FALCON),
 that employs metabolic network reconstructions along with expression
data to estimate fluxes. In order to use such a method, accurate
measures of enzyme complex abundance are needed, so we first
present a new algorithm that addresses quantification of complex
abundance. Our extensions to prior techniques include the
capability to work with large models and significantly improved
run-time performance even for smaller models, an improved analysis of
enzyme complex formation logic, the ability to handle very large enzyme
complex rules that may incorporate multiple isoforms, and depending on
the model constraints, either maintained or significantly improved
correlation with experimentally measured fluxes.
}

\newcommand{\falconAbstractAvail}{%
FALCON has been implemented in MATLAB and ATS, and can be downloaded
from: \url{https://github.com/bbarker/FALCON}. ATS is not required to
compile the software, as intermediate C source code is available, and
binaries are provided for Linux x86-64 systems. FALCON requires use
of the COBRA Toolbox, also implemented in MATLAB.
}

\captionsetup{labelfont=bf}


% Take care of potentially defined variables:

% \newcommand[1]{\identifndef}{
%   \if\isdef\csname{#1}
%     {}
%   \else
%     \newcommand{\{#1}}[1]{\{#1}}
%   \fi
% }

% \identifndef{processtable}