\documentclass[letterpaper]{article}
\usepackage[affil-it]{authblk}

% \makeatletter
% %
% \def\maketitle
%   {{\@flushglue=.25\textwidth minus.25\textwidth\z@skip
%     \hyphenpenalty\@M
%     \let\old@date\@date
%     \def\@date{\mbox{}\hskip\@flushglue\old@date\hskip\@flushglue\mbox{}\par}%hmm
%     \renewenvironment{tabular}[2][]{\par}
%                                    {\par}%
%   \AB@maketitle}}
% %
% \makeatother


\usepackage{amsmath}
\usepackage{amsthm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                                       %
% change this line for portable code:  %
\newcommand*{\commonDir}{./common/}    %
\input{\commonDir preambleCommon}      %
                                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{natbib}
\bibliographystyle{natbib}


\graphicspath{{./falcon/figures/}}

\usepackage{url}
\usepackage{fullpage}
\usepackage{setspace}

\captionsetup[ruled]{labelsep=period}
\makeatletter
\@addtoreset{algorithm}{section}% algorithm counter resets every section
\makeatother
\renewcommand{\thealgorithm}{\thesection.\arabic{algorithm}}% Algorithm # is <section>.<algorithm>


\title{A robust and efficient method for estimating enzyme
complex abundance and metabolic flux from expression data}

\author[1]{Brandon Barker%
  \thanks{Electronic address: \texttt{brandon.barker@cornell.edu}; Corresponding author}}

\author[2]{Narayanan Sadagopan}
%  \thanks{Electronic address: \texttt{narayanan.sadagopan@umassmed.edu}}}

\author[3]{Yiping Wang}
%  \thanks{Electronic address: \texttt{yw595@cornell.edu}}}

\author[4,5]{Kieran Smallbone}
%  \thanks{Electronic address: \texttt{kieran.smallbone@manchester.ac.uk}}}

\author[3,6,7]{Christopher R. Myers}
%  \thanks{Electronic address: \texttt{c.myers@cornell.edu}}}

\author[8]{Hongwei Xi}
% \author[8]{Hongwei Xi%
%   \thanks{Electronic address: \texttt{hwxi@bu.edu}}}

\author[3,9]{Jason W. Locasale%
  \thanks{Electronic address: \texttt{locasale@cornell.edu}; Corresponding author}}

\author[3,9]{Zhenglong Gu%
  \thanks{Electronic address: \texttt{zg27@cornell.edu}; Corresponding author}}

\affil[1]{Center for Advanced Computing,
  Cornell University, 534 Rhodes Hall, Ithaca, NY, USA.}
%\affil[2]{Department of Biological Statistics and Computational Biology,
%    Cornell University, 1198 Comstock Hall, Ithaca, NY, USA.}
\affil[2]{Program of Bioinformatics and Integrative Biology,
  University of Massachusetts Medical School,
  55 Lake Avenue North, Worcester MA 01655}
\affil[3]{Tri-Institutional Training Program in Computational
  Biology and Medicine, 1300 York Avenue, Box 194, New York, NY, USA.}
\affil[4]{School of Computer Science, The University of Manchester,
  Manchester, UK.}
\affil[5]{Manchester Center for Integrative Systems Biology, 
  The University of Manchester, \mbox{Manchester, UK.}}
\affil[6]{Laboratory of Atomic and Solid State Physics, Cornell
University, Ithaca, NY, USA.}
\affil[7]{Institute of Biotechnology, Cornell University, Ithaca, NY,
USA.}
\affil[8]{Department of Computer Science, Boston University,
  111 Cummington Street, Boston, MA, USA.}
\affil[9]{Division of Nutritional Sciences, Cornell University,
  Savage Hall, Ithaca, NY, USA.}


\date{\today}

%%%%%%%%%%%%%%%%%%
\begin{document}%%
%%%%%%%%%%%%%%%%%%

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                                       %
\newboolean{thesisStyle}               %
\setboolean{thesisStyle}{true}         %
                                       %
\input{\commonDir documentHeadCommon}  %
                                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
\falconAbstractMotivation
\falconAbstractResults
\falconAbstractAvail
\end{abstract}

% Redefine suppOrApp
\def\suppOrApp{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                          %
\input{\commonDir falcon}  %
%                          %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                    %
\input{\commonDir falcon_appendix}   %
%                                    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\bibliography{library}
\end{document}
