\documentclass{bioinfo}
\usepackage{xr}
\externaldocument{supplement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                                       %
% change this line for portable code:  %
\newcommand*{\commonDir}{../common/}   %
\input{\commonDir preambleCommon}      %
                                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{natbib}

\graphicspath{{./figures/}}

\usepackage{url}

%Along with the tikz arrows and such, these seem to conflict with the bionfo header:
%%% \usepackage{amsmath}
%%% \usepackage{hyperref}

%\usepackage[amsmath,amsthm,thmmarks]{ntheorem}


% \renewcommand{\caption}[1]{\singlespacing\hangcaption{#1}\normalspacing}


\copyrightyear{2013}
\pubyear{2013}


%%%%%%%%%%%%%%%%%%
\begin{document}%%
%%%%%%%%%%%%%%%%%%

\renewcommand{\figurename}{\textbf{Fig.}}
\captionsetup{belowskip=0pt}

\newboolean{thesisStyle}
\setboolean{thesisStyle}{false}

\firstpage{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
                                       %
\input{\commonDir documentHeadCommon}  %
                                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% A systematic exploration of enzyme complex rules and metabolic networks yields fast and
% robust algorithms for estimating metabolic flux and enzyme complex abundance.

% Metabolic flux and enzyme complex abundance are efficiently and robustly
% estimated by systematic inspection of GPR rule sand enzyme complex abundance.

% Systematic consideration of gene-rules and enzyme complex relationships enables 
% quantification of fluxes and enzyme complexes.

% Fast and systematic algorithms estimate enzyme complex abundance
% and flux from expression data.


\title[FALCON]{A robust and efficient method for\\ 
estimating enzyme complex abundance\\ 
and metabolic flux from expression data}
%    \author[Barker \textit{et~al.}]{Brandon Barker et al.$^{1,2}$\footnote{to whom correspondence should be addressed}}
  \author[Barker \textit{et~al.}]{Brandon Barker\,$^{1,2}$\footnote{to whom correspondence should be addressed},
    Narayanan Sadagopan\,$^{3}$, Yiping Wang\,$^{1,2}$, \\
    Kieran Smallbone\,$^{4,5}$, Christopher R. Myers\,$^{1,6,7}$, Hongwei Xi\,$^{8}$, \\ 
    Jason W. Locasale\,$^{1,9*}$, and Zhenglong Gu\,$^{1,9*}$\\
 }

\address{$^{1}$Tri-Institutional Training Program in Computational
  Biology and Medicine, 1300 York Avenue, Box 194, New York, NY, USA.
  $^{2}$Department of Biological Statistics and Computational Biology, 
    Cornell University, 1198 Comstock Hall, Ithaca, NY, USA.
  $^{3}$College of Engineering, Cornell University, Carpenter Hall, 
    Ithaca, NY, USA.
  $^{4}$ School of Computer Science, The University of Manchester, Manchester, UK.
  $^{5}$Manchester Center for Integrative Systems Biology, The University of Manchester, Manchester, UK.
  $^{6}$Laboratory of Atomic and Solid State Physics, Cornell University, Ithaca, NY, USA.
  $^{7}$Institute of Biotechnology, Cornell University, Ithaca, NY, USA.
  $^{8}$Department of Computer Science, Boston University,
    111 Cummington Street, Boston, MA, USA.
  $^{9}$Division of Nutritional Sciences, Cornell University, 
    Savage Hall, Ithaca, NY, USA.
  }
\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: XXXXXXX}

\maketitle

\begin{abstract}

\section{Motivation:}
\falconAbstractMotivation

\section{Results:}
\falconAbstractResults

\section{Availability:}
\falconAbstractAvail

\section{Supporting information:}
Supporting information is available at \emph{Bioinformatics}
online.

\section{Contact:} \href{beb82@cornell.edu}{beb82@cornell.edu},
\href{locasale@cornell.edu}{locasale@cornell.edu},
\href{zg27@cornell.edu}{zg27@cornell.edu}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                              %
\input{\commonDir falcon.tex}  %
%                              %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgement}

We thank Neil Swainston for his previous work on these research issues
and many helpful discussions about considerations in various genome
scale models of metabolism.  We also thank Michael Stillman for
general discussions about modeling and for reading the paper. Finally
we thank Alex Shestov, Lei Huang, and Eli Bogart for many helpful
discussions about metabolism.

\paragraph{Funding\textcolon}
BEB and YW are supported by the Tri-Institutional Training Program in Computational Biology and Medicine.
BEB and ZG are thankful for support from NSF grant MCB-1243588.
KS is grateful for the financial support of the EU FP7 project BioPreDyn (grant 289434).
CRM acknowledges support from NSF Grant IOS-1127017.
JWL thanks the NIH for support through grants R00 CA168997 and R01 AI110613.


%\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
%\bibliographystyle{plain}
%
%\bibliography{Document}


\bibliography{dissertation}

\end{document}
